Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics

### 2020-01-12 paper/paper.Rmd
methods section - loom for XXX to revise

@@

The lineup protocol allows a known model to be imposed on the null data set plots.
The underlying spatial correlation model was created to provide spatial autocorrelation between neighboring areas using the longitude and latitude values for the Statistical Areas.

$$z = 1$$
$$locations = longitude + latitude$$

@@

## Experiment procudure and data collection

@@

## Experimental design

@@

### visualizations
Side-by-side dot plots were made of accuracy (efficiency) against map type, facetted by trend model type.

@@

## Limitations of the data collection

A pilot study was conducted to determine whether the lineups were appropriate. All lineups in the pilot study were deemed viable as at least one participant detected the real data plot in each lineup.
The demographics of the participants showed a skew towards male participants. 
The randomness of the group allocation also resulted in more participants being allocated to Group B. Due to the allocation of lineup displays the participants all saw six Choropleth displays and six Hexagon Tile Map displays.

@@

@@

>>>>

Australia is used for the study, with Statistical Area 2 (SA2) (XXX REF) as the geographic units. The results shuold apply broadly to any other geographic area of interest.

@@

Null data needs to be data with no (interesting) structure. In most scenarios, permutation is the main approach for generating null plots. It is used to break association between variables, while maintaining marginal distributions. This is too simple for spatial data. In spatial data, a key feature is smoothness over the landscape corresponding spatial dependence. To do something simple, like permute the values relative to the goegraphic location would produce null plots which are too chaotic, and the data plot will be recognisable for its smoothness rather than any structure of interest.

For spatial data, null data is staionary data, where the mean, variance and spatial dependence are constant over the geographic units. Stationary data is specified by a variogram model (XXX REF). Simulating from a variogram model, where the spatial dependence is specified, generates the stationary spatial data used for the null plots. A Gaussian The parameters for the model were sill=1, range=0.3 with the variance generated by a standard normal distribution (XXX What parameters were used?).

The R package `gstat` (XXX REF) was used to simulate 144 null sets, 12 data sets for each plot in a lineup, and 12 sets for 12 lineups.

```{r eval=FALSE, echo=FALSE}
var.g.dummy <- gstat(formula = z ~ 1, 
                     locations = ~ longitude + latitude, 
                     dummy = T, beta = 1, model = vgm(psill = 1, model = "Gau", range = 0.3),
                     nmax = 12)
```

<!-- Spatially dependent data sets were simulated using the variogram model on the centroids of each geographic unit. Twelve simulations from the variogram model were created for each of the twelve lineups.

The underlying spatial correlation model was created to provide spatial autocorrelation between neighboring areas using the longitude and latitude values for the Statistical Areas.

$$z = 1$$
$$locations = longitude + latitude$$
-->

## Generating lineups

XXX Fill in this structure - see commented text

<!-- Discuss simulating the trend data -->

<!-- Applying smoothing -->

<!-- Scaling data within a lineup -->

<!-- Location of data plot -->

@@

For each lineup these statistics were calculated:

## Web app collecting responses

XXX Describe the details of the taipan app used to collect responses

XXX Sentence on pilot study which allowed us to estimate the effect size, and thus decide on number of participants to collect responses from

## Demographic data collection

@@

<!--
## Experimental design

@@

 - accuracy: the proportion of subjects who detected the data plot
 - efficiency: average time taken to respond
-->

## Analysis

### Descriptive statistics

Mean, standard deviations for lineups, by participant, ...

Side-by-side dot plots were made of accuracy (efficiency) against map type, facetted by trend model type.

@@

### Data cleaning

<!-- A pilot study was conducted to determine whether the lineups were appropriate. All lineups in the pilot study were deemed viable as at least one participant detected the real data plot in each lineup.
The demographics of the participants showed a skew towards male participants. 
The randomness of the group allocation also resulted in more participants being allocated to Group B. Due to the allocation of lineup displays the participants all saw six Choropleth displays and six Hexagon Tile Map displays.
-->

@@

Acknowledgements
==============

@@

All code to generate the data, of the web app that collected participants responses, conduct the analysis, and produce this manuscript are available on XXX.

The supplementary material contains:

- Additonal analysis of the experimental results
- Training materials used for the participants
- 24 lineups used in the experiment
- 12 data sets used to coonstruct the lineups

### 2020-01-12 paper/paper.tex
methods section - loom for XXX to revise

The lineup protocol allows a known model to be imposed on the null data set plots.
The underlying spatial correlation model was created to provide spatial autocorrelation between neighboring areas using the longitude and latitude values for the Statistical Areas.

\[z = 1\]
\[locations = longitude + latitude\]

@@

\hypertarget{experiment-procudure-and-data-collection}{%
\subsection{Experiment procudure and data collection}\label{experiment-procudure-and-data-collection}}

@@

\hypertarget{experimental-design}{%
\subsection{Experimental design}\label{experimental-design}}

The choropleth map was used as the comparative visualization for presenting the lineups {[}\protect\hyperlink{ref-VVSIALM}{15}{]} as this is the common display for spatial cancer data.
Geographic distributions usually have some degree of spatial autcorrelation between neighbors.
This feature was incorporated in all maps shown in the lineup displays, the map that contained the trend feature shown in only one set of data was also affected by spatial autocorrelation.
A line up protocol was implemented to arrange \(N\) maps in each display.
A reasonable amount of null plots \(N-1\) in the lineup was chosen to ensure the real data map was well hidden. A reasonable number of plots to show in each lineup, \(N = 12\) was chosen to not overwhelm participants due to the detailed choropleth maps of Australian SA3 areas.

The hypotheses for each lineup are
\(H_0\) : All plots look the same
\(H_a\) : One plot looks different to the other plots

The same data was visualized on a choropleth map, and on a hexagon tile map, however in this study the participants only saw one of these two displays. The accuracy and times taken were aggregated for Groups and.
Comparing the results of participants who see the choropleth to those who see a hexagon tile map will show that population related distributions are spotted more frequently in a hexagon tile map display.

Let \(n\) be the number of independent observers and \(x_i\) the
number of observers who picked plot \(i\), \(i = \{1,...,m\}\)

Then \(x_i, x_2, ..., x_m\) follows a multinomial
distribution\(Mult_{\pi_1, \pi_2, ...., \pi_m}(x_i, x_2, ..., x_m)\) with
\(\sum_i \pi_i = 1\), where \(\pi_i\) is the probability that plot \(i\) is picked by an
observer, which we can estimate as \(\hat{\pi}_i = x_i/n\).
The researchers compared the length of time taken, and the accuracy of the participants choices.
The power of a lineup can therefore be estimated as the ratio of correct
identifications \(x\) out of \(n\) viewings.

\hypertarget{the-methods-of-data-analysis-used}{%
\subsection{The methods of data analysis used}\label{the-methods-of-data-analysis-used}}

The data analysis methods used in order to analyse and collate the results included downloading the survey submissions and opening them into the analysis software R {[}\protect\hyperlink{ref-RCore}{16}{]}.

For each of the 12 lineup displays the researchers calculated:

\hypertarget{visualizations}{%
\subsubsection{visualizations}\label{visualizations}}

@@

The likelihood of detecting the data plot in the lineup can be modelled using a linear mixed effects model.
The R {[}\protect\hyperlink{ref-RCore}{16}{]} \texttt{glmer()} function in the {[}\protect\hyperlink{ref-lme4}{17}{]} package implements generalised linear mixed effect models. The model used includes the two main effects map type and trend model, which gives the fixed effects model to be:

@@

Computation will be done using R {[}\protect\hyperlink{ref-RCore}{16}{]}, with the \texttt{lme4} package {[}\protect\hyperlink{ref-lme4}{17}{]}.

\hypertarget{limitations-of-the-data-collection}{%
\subsection{Limitations of the data collection}\label{limitations-of-the-data-collection}}

A pilot study was conducted to determine whether the lineups were appropriate. All lineups in the pilot study were deemed viable as at least one participant detected the real data plot in each lineup.
The demographics of the participants showed a skew towards male participants.
The randomness of the group allocation also resulted in more participants being allocated to Group B. Due to the allocation of lineup displays the participants all saw six Choropleth displays and six Hexagon Tile Map displays.

@@

@@

@@

\hypertarget{acknowledgment-1}{%
\section{Acknowledgment}\label{acknowledgment-1}}

@@

The analysis of the work was completed in R {[}\protect\hyperlink{ref-RCore}{16}{]} with the use of the following packages:

\begin{itemize}
\tightlist
\item
  For document creation: rmarkdown {[}\protect\hyperlink{ref-rmarkdown}{18}{]}, rticles {[}\protect\hyperlink{ref-rticles}{19}{]}, knitr {[}\protect\hyperlink{ref-knitr}{20}{]}.
\item
  For lineup creation and data analysis: tidyverse {[}\protect\hyperlink{ref-tidyverse}{21}{]}, nullabor {[}\protect\hyperlink{ref-nullabor}{22}{]}, ggthemes {[}\protect\hyperlink{ref-ggthemes}{23}{]}, RColorBrewer {[}\protect\hyperlink{ref-RColorBrewer}{24}{]}.
\item
  For image displays: cowplot {[}\protect\hyperlink{ref-cowplot}{25}{]}, png {[}\protect\hyperlink{ref-png}{26}{]}, grid {[}\protect\hyperlink{ref-grid}{27}{]}.
\item
  For modeling and presentation of models: lme4 {[}\protect\hyperlink{ref-lme4}{17}{]}, kableExtra {[}\protect\hyperlink{ref-kableExtra}{28}{]}.
\end{itemize}

@@

\leavevmode\hypertarget{ref-VVSIALM}{}%
15. Majumder M, Hofmann H, Cook D (2013) Validation of visual statistical inference, applied to linear models. Journal of the American Statistical Association 108:942--956

\leavevmode\hypertarget{ref-RCore}{}%
16. R Core Team (2019) R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria

\leavevmode\hypertarget{ref-lme4}{}%
17. Bates D, Mächler M, Bolker B, Walker S (2015) Fitting linear mixed-effects models using lme4. Journal of Statistical Software 67:1--48

\leavevmode\hypertarget{ref-rmarkdown}{}%
18. Xie Y, Allaire JJ, Grolemund G (2018) R markdown: The definitive guide. Chapman; Hall/CRC, Boca Raton, Florida

\leavevmode\hypertarget{ref-rticles}{}%
19. Allaire J, Xie Y, R Foundation, et al (2019) Rticles: Article formats for r markdown.

\leavevmode\hypertarget{ref-knitr}{}%
20. Xie Y (2014) Knitr: A comprehensive tool for reproducible research in R. Implementing reproducible computational research

\leavevmode\hypertarget{ref-tidyverse}{}%
21. Wickham H, Averick M, Bryan J, et al (2019) Welcome to the tidyverse. Journal of Open Source Software 4:1686

\leavevmode\hypertarget{ref-nullabor}{}%
22. Wickham H, Chowdhury NR, Cook D, Hofmann H (2018) Nullabor: Tools for graphical inference.

\leavevmode\hypertarget{ref-ggthemes}{}%
23. Arnold JB (2019) Ggthemes: Extra themes, scales and geoms for 'ggplot2'.

\leavevmode\hypertarget{ref-RColorBrewer}{}%
24. Neuwirth E (2014) RColorBrewer: ColorBrewer palettes.

\leavevmode\hypertarget{ref-cowplot}{}%
25. Wilke CO (2019) Cowplot: Streamlined plot theme and plot annotations for 'ggplot2'.

\leavevmode\hypertarget{ref-png}{}%
26. Urbanek S (2013) Png: Read and write png images.

\leavevmode\hypertarget{ref-grid}{}%
27. R Core Team (2019) R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria

\leavevmode\hypertarget{ref-kableExtra}{}%
28. Zhu H (2019) KableExtra: Construct complex table with 'kable' and pipe syntax.

>>>>

Null data needs to be data with no (interesting) structure. In most scenarios, permutation is the main approach for generating null plots. It is used to break association between variables, while maintaining marginal distributions. This is too simple for spatial data. In spatial data, a key feature is smoothness over the landscape corresponding spatial dependence. To do something simple, like permute the values relative to the goegraphic location would produce null plots which are too chaotic, and the data plot will be recognisable for its smoothness rather than any structure of interest.

For spatial data, null data is staionary data, where the mean, variance and spatial dependence are constant over the geographic units. Stationary data is specified by a variogram model (XXX REF). Simulating from a variogram model, where the spatial dependence is specified, generates the stationary spatial data used for the null plots. A Gaussian The parameters for the model were sill=1, range=0.3 with the variance generated by a standard normal distribution (XXX What parameters were used?).

The R package \texttt{gstat} (XXX REF) was used to simulate 144 null sets, 12 data sets for each plot in a lineup, and 12 sets for 12 lineups.

\hypertarget{generating-lineups}{%
\subsection{Generating lineups}\label{generating-lineups}}

XXX Fill in this structure - see commented text

@@

For each lineup these statistics were calculated:

\hypertarget{web-app-collecting-responses}{%
\subsection{Web app collecting responses}\label{web-app-collecting-responses}}

XXX Describe the details of the taipan app used to collect responses

XXX Sentence on pilot study which allowed us to estimate the effect size, and thus decide on number of participants to collect responses from

\hypertarget{demographic-data-collection}{%
\subsection{Demographic data collection}\label{demographic-data-collection}}

@@

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

\hypertarget{descriptive-statistics}{%
\subsubsection{Descriptive statistics}\label{descriptive-statistics}}

Mean, standard deviations for lineups, by participant, \ldots{}

@@

The likelihood of detecting the data plot in the lineup can be modelled using a linear mixed effects model.
The R {[}\protect\hyperlink{ref-RCore}{15}{]} \texttt{glmer()} function in the {[}\protect\hyperlink{ref-lme4}{16}{]} package implements generalised linear mixed effect models. The model used includes the two main effects map type and trend model, which gives the fixed effects model to be:

@@

Computation will be done using R {[}\protect\hyperlink{ref-RCore}{15}{]}, with the \texttt{lme4} package {[}\protect\hyperlink{ref-lme4}{16}{]}.

\hypertarget{data-cleaning}{%
\subsubsection{Data cleaning}\label{data-cleaning}}

@@

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

@@

All code to generate the data, of the web app that collected participants responses, conduct the analysis, and produce this manuscript are available on XXX.

The supplementary material contains:

\begin{itemize}
\tightlist
\item
  Additonal analysis of the experimental results
\item
  Training materials used for the participants
\item
  24 lineups used in the experiment
\item
  12 data sets used to coonstruct the lineups
\end{itemize}

@@

@@

The analysis of the work was completed in R {[}\protect\hyperlink{ref-RCore}{15}{]} with the use of the following packages:

\begin{itemize}
\tightlist
\item
  For document creation: rmarkdown {[}\protect\hyperlink{ref-rmarkdown}{17}{]}, rticles {[}\protect\hyperlink{ref-rticles}{18}{]}, knitr {[}\protect\hyperlink{ref-knitr}{19}{]}.
\item
  For lineup creation and data analysis: tidyverse {[}\protect\hyperlink{ref-tidyverse}{20}{]}, nullabor {[}\protect\hyperlink{ref-nullabor}{21}{]}, ggthemes {[}\protect\hyperlink{ref-ggthemes}{22}{]}, RColorBrewer {[}\protect\hyperlink{ref-RColorBrewer}{23}{]}.
\item
  For image displays: cowplot {[}\protect\hyperlink{ref-cowplot}{24}{]}, png {[}\protect\hyperlink{ref-png}{25}{]}, grid {[}\protect\hyperlink{ref-grid}{26}{]}.
\item
  For modeling and presentation of models: lme4 {[}\protect\hyperlink{ref-lme4}{16}{]}, kableExtra {[}\protect\hyperlink{ref-kableExtra}{27}{]}.
\end{itemize}

@@

\leavevmode\hypertarget{ref-RCore}{}%
15. R Core Team (2019) R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria

\leavevmode\hypertarget{ref-lme4}{}%
16. Bates D, Mächler M, Bolker B, Walker S (2015) Fitting linear mixed-effects models using lme4. Journal of Statistical Software 67:1--48

\leavevmode\hypertarget{ref-rmarkdown}{}%
17. Xie Y, Allaire JJ, Grolemund G (2018) R markdown: The definitive guide. Chapman; Hall/CRC, Boca Raton, Florida

\leavevmode\hypertarget{ref-rticles}{}%
18. Allaire J, Xie Y, R Foundation, et al (2019) Rticles: Article formats for r markdown.

\leavevmode\hypertarget{ref-knitr}{}%
19. Xie Y (2014) Knitr: A comprehensive tool for reproducible research in R. Implementing reproducible computational research

\leavevmode\hypertarget{ref-tidyverse}{}%
20. Wickham H, Averick M, Bryan J, et al (2019) Welcome to the tidyverse. Journal of Open Source Software 4:1686

\leavevmode\hypertarget{ref-nullabor}{}%
21. Wickham H, Chowdhury NR, Cook D, Hofmann H (2018) Nullabor: Tools for graphical inference.

\leavevmode\hypertarget{ref-ggthemes}{}%
22. Arnold JB (2019) Ggthemes: Extra themes, scales and geoms for 'ggplot2'.

\leavevmode\hypertarget{ref-RColorBrewer}{}%
23. Neuwirth E (2014) RColorBrewer: ColorBrewer palettes.

\leavevmode\hypertarget{ref-cowplot}{}%
24. Wilke CO (2019) Cowplot: Streamlined plot theme and plot annotations for 'ggplot2'.

\leavevmode\hypertarget{ref-png}{}%
25. Urbanek S (2013) Png: Read and write png images.

\leavevmode\hypertarget{ref-grid}{}%
26. R Core Team (2019) R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria

\leavevmode\hypertarget{ref-kableExtra}{}%
27. Zhu H (2019) KableExtra: Construct complex table with 'kable' and pipe syntax.
